VSTE - Comparing Methods Visualization

### 2022-12-08 TestingPipeline/Scripts/comparing_methods_visualization.Rmd
Most updated rules.

  if (include_vibrant) {
    keep_score[input_seqs$vibrant_quality=="complete circular"] <- keep_score[input_seqs$vibrant_quality=="complete circular"] + 1
    keep_score[input_seqs$vibrant_quality=="high quality draft"] <- keep_score[input_seqs$vibrant_quality=="high quality draft"] + 1
    keep_score[input_seqs$vibrant_quality=="medium quality draft"] <- keep_score[input_seqs$vibrant_quality=="medium quality draft"] + 1
    keep_score[input_seqs$vibrant_quality=="low quality draft" & input_seqs$provirus] <- keep_score[input_seqs$vibrant_quality=="low quality draft" & input_seqs$provirus] + 0.5
  }

  if (include_virsorter2) {
    keep_score[input_seqs$viral>=50] <- keep_score[input_seqs$viral>=50] + 0.5
    keep_score[input_seqs$viral>=95] <- keep_score[input_seqs$viral>=95] + 0.5
    keep_score[input_seqs$RNA>=0.9] <- keep_score[input_seqs$RNA>=0.9] + 1
    keep_score[input_seqs$lavidaviridae>=0.9] <- keep_score[input_seqs$lavidaviridae>=0.9] + 1
    keep_score[input_seqs$NCLDV>=0.9] <- keep_score[input_seqs$NCLDV>=0.9] + 1
  }

  if (include_virsorter) {
    keep_score[input_seqs$category==1] <- keep_score[input_seqs$category==1] + 1
    keep_score[input_seqs$category==2] <- keep_score[input_seqs$category==2] + 0.5
    keep_score[input_seqs$category==4] <- keep_score[input_seqs$category==4] + 1
    keep_score[input_seqs$category==5] <- keep_score[input_seqs$category==5] + 0.5
  }

  if (include_deepvirfinder) {
    keep_score[input_seqs$score>=0.7 & input_seqs$checkv_length<20000] <- keep_score[input_seqs$score>=0.7 & input_seqs$checkv_length<20000] + 0.5
    keep_score[input_seqs$score>=0.9 & input_seqs$checkv_length<20000] <- keep_score[input_seqs$score>=0.9 & input_seqs$checkv_length<20000] + 0.5
  }

  if (include_tuning_viral) {
    keep_score[input_seqs$Kaiju_Viral=="Viruses"] <- keep_score[input_seqs$Kaiju_Viral=="Viruses"] + 0.5
    keep_score[input_seqs$hallmark>2] <- keep_score[input_seqs$hallmark>2] + 1
    keep_score[input_seqs$percent_unknown>=75 & input_seqs$checkv_length<50000] <- keep_score[input_seqs$percent_unknown>=75 & input_seqs$checkv_length<50000] + 0.5
    keep_score[input_seqs$percent_viral>=50] <- keep_score[input_seqs$percent_viral>=50] + 0.5
  }

  if (include_tuning_not_viral) {
    keep_score[input_seqs$Kaiju_Viral=="cellular organisms"] <- keep_score[input_seqs$Kaiju_Viral=="cellular organisms"] - 0.5
    keep_score[input_seqs$checkv_host_genes>50 & !input_seqs$provirus] <- keep_score[input_seqs$checkv_host_genes>50 & !input_seqs$provirus] - 1
    keep_score[input_seqs$checkv_viral_genes==0 & input_seqs$checkv_host_genes>=1] <- keep_score[input_seqs$checkv_viral_genes==0 & input_seqs$checkv_host_genes>=1] - 1
    keep_score[((input_seqs$checkv_viral_genes*3) <= input_seqs$checkv_host_genes) & !input_seqs$provirus] <- keep_score[((input_seqs$checkv_viral_genes*3) <= input_seqs$checkv_host_genes) & !input_seqs$provirus] - 1 # consider accounting for RNA viruses
    keep_score[input_seqs$checkv_length>500000 & input_seqs$hallmark<=1] <- keep_score[input_seqs$checkv_length>500000 & input_seqs$hallmark<=1] - 1
    keep_score[input_seqs$checkv_completeness<=75 & input_seqs$checkv_length<=5000] <- keep_score[input_seqs$checkv_completeness<=75 & input_seqs$checkv_length<=5000] - 0.5 # helped with protist contamination
  }

@@

@@

```{r}
hist(prop_high_viruses_long$proportion,
     )
```

>>>>

  if (include_vibrant) {
    keep_score[input_seqs$vibrant_quality=="high quality draft"] <- keep_score[input_seqs$vibrant_quality=="high quality draft"] + 1
    keep_score[input_seqs$vibrant_quality=="medium quality draft"] <- keep_score[input_seqs$vibrant_quality=="medium quality draft"] + 1
    keep_score[input_seqs$vibrant_quality=="low quality draft" & input_seqs$provirus] <- keep_score[input_seqs$vibrant_quality=="low quality draft" & input_seqs$provirus] + 0.5
  }

  if (include_virsorter2) {
    keep_score[input_seqs$max_score>=0.50] <- keep_score[input_seqs$max_score>=0.50] + 0.5
    keep_score[input_seqs$max_score>=0.95] <- keep_score[input_seqs$max_score>=0.95] + 0.5  
  }

  if (include_virsorter) {
    keep_score[input_seqs$category==1] <- keep_score[input_seqs$category==1] + 1
    keep_score[input_seqs$category==2] <- keep_score[input_seqs$category==2] + 0.5
    keep_score[input_seqs$category==3] <- keep_score[input_seqs$category==3] + 0.5
    keep_score[input_seqs$category==4] <- keep_score[input_seqs$category==4] + 1
    keep_score[input_seqs$category==5] <- keep_score[input_seqs$category==5] + 0.5 
    keep_score[input_seqs$category==6] <- keep_score[input_seqs$category==6] + 0.5 
  }

  if (include_deepvirfinder) {
     # add if DVF calls viral
    keep_score[(input_seqs$score>=0.9 & input_seqs$pvalue<=0.05) & input_seqs$checkv_length<20000] <- keep_score[input_seqs$score>=0.9 & input_seqs$checkv_length<20000] + 0.5
    keep_score[(input_seqs$score>=0.7 & input_seqs$pvalue<=0.05) & input_seqs$checkv_length<20000] <- keep_score[input_seqs$score>=0.7 & input_seqs$checkv_length<20000] + 0.5
  }

  if (include_tuning_viral) {
  # tuning addition
    keep_score[input_seqs$hallmark>2] <- keep_score[input_seqs$hallmark>2] + 1
    keep_score[input_seqs$Kaiju_Viral=="Viruses"] <- keep_score[input_seqs$Kaiju_Viral=="Viruses"] + 1
    #note: had tried pulling out this rule, but better recall without sacrificing precision with it in
    #note: having kaiju removal rule didn't help
    keep_score[input_seqs$percent_unknown>=75 & input_seqs$checkv_length<50000] <- keep_score[input_seqs$percent_unknown>=75 & input_seqs$checkv_length<50000] + 0.5
    keep_score[input_seqs$viral>=50 | input_seqs$percent_viral>=50] <- keep_score[input_seqs$viral>=50 | input_seqs$percent_viral>=50] + 0.5 
    keep_score[(input_seqs$checkv_completeness>=75 | input_seqs$vibrant_quality=="complete circular") & (input_seqs$viral>=50 | input_seqs$percent_viral>=50)] <- keep_score[(input_seqs$checkv_completeness>=75 | input_seqs$vibrant_quality=="complete circular") & (input_seqs$viral>=50 | input_seqs$percent_viral>=50)] + 0.5 
  }

  if (include_tuning_not_viral) {
    # tuning removal 
    keep_score[input_seqs$checkv_host_genes>50 & !input_seqs$provirus] <- 0
    keep_score[input_seqs$checkv_viral_genes==0 & input_seqs$checkv_host_genes>=1] <- keep_score[input_seqs$checkv_viral_genes==0 & input_seqs$checkv_host_genes>=1] - 1
    keep_score[((input_seqs$checkv_viral_genes*3) <= input_seqs$checkv_host_genes) & !input_seqs$provirus] <- keep_score[((input_seqs$checkv_viral_genes*3) <= input_seqs$checkv_host_genes) & !input_seqs$provirus] - 1 
    keep_score[input_seqs$checkv_length>500000 & input_seqs$hallmark<=1] <- 0
    keep_score[(input_seqs$checkv_completeness<=75) & input_seqs$checkv_length<=5000] <- keep_score[(input_seqs$checkv_completeness<=75) & input_seqs$checkv_length<=5000] - 0.5 
    # remove if DVF calls it not viral
    keep_score[input_seqs$score<=0.7 & input_seqs$pvalue<=0.05] <- keep_score[input_seqs$score<=0.7 & input_seqs$pvalue<=0.05] - 1

  }

@@

# overlap in proportion of contigs for tool sets

get number of tools and number of rules for each combination
```{r}
ann_col$numrules <- rowSums(ann_col[,1:6])
ann_col$numtools <- ann_col$numrules

for (i in 1:nrow(ann_col)) {
  if (ann_col$tv[i]==1) {
    ann_col$numtools[i] <- ann_col$numtools[i] + 3
  }

  if (ann_col$tnv[i]==1) {
    ann_col$numtools[i] <- ann_col$numtools[i] + 5
  }
}

ann_col$numtools[ann_col$numtools>6] <- 6

ann_col$tool <- rownames(ann_col)
```

prop_high_viruses_long <- as.data.frame(prop_high_viruses_long) %>% 
  pivot_longer(!tool,
               values_to="proportion",
               names_to="tool2") %>%
  inner_join(ann_col, by=c("tool"))

```{r}
ggplot(prop_high_viruses_long[prop_high_viruses_long$vs2==1,], aes(x=as.factor(numrules), y=proportion, 
                                  color=as.factor(numrules), fill=as.factor(numrules))) +
  geom_boxplot() +
  #geom_point() +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Number of Rules") +
  ylab("Proportion of Contigs in Common") +
  scale_fill_manual(name="",
                     values = alpha(rev(pal(6)), 0.3)) +
  scale_color_manual(name="",
                     values = alpha(rev(pal(6)), 0.7))

ggplot(prop_high_viruses_long[prop_high_viruses_long$vs==1,], aes(x=as.factor(numrules), y=proportion, 
                                  color=as.factor(numrules), fill=as.factor(numrules))) +
  geom_boxplot() +
  #geom_point() +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Number of Rules") +
  ylab("Proportion of Contigs in Common") +
  scale_fill_manual(name="",
                     values = alpha(rev(pal(6)), 0.3)) +
  scale_color_manual(name="",
                     values = alpha(rev(pal(6)), 0.7))

ggplot(prop_high_viruses_long[prop_high_viruses_long$vb==1,], aes(x=as.factor(numrules), y=proportion, 
                                  color=as.factor(numrules), fill=as.factor(numrules))) +
  geom_boxplot() +
  #geom_point() +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Number of Rules") +
  ylab("Proportion of Contigs in Common") +
  scale_fill_manual(name="",
                     values = alpha(rev(pal(6)), 0.3)) +
  scale_color_manual(name="",
                     values = alpha(rev(pal(6)), 0.7))

ggplot(prop_high_viruses_long[prop_high_viruses_long$dvf==1,], aes(x=as.factor(numrules), y=proportion, 
                                  color=as.factor(numrules), fill=as.factor(numrules))) +
  geom_boxplot() +
  #geom_point() +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Number of Rules") +
  ylab("Proportion of Contigs in Common") +
  scale_fill_manual(name="",
                     values = alpha(rev(pal(6)), 0.3)) +
  scale_color_manual(name="",
                     values = alpha(rev(pal(6)), 0.7))
```

```{r}
table(prop_high_viruses_long$numrules[prop_high_viruses_long$dvf==1])

```{r}
ggplot(prop_high_viruses_long[prop_high_viruses_long$vs2==1,], aes(x=as.factor(numtools), y=proportion, 
                                  color=as.factor(numtools), fill=as.factor(numtools))) +
  geom_boxplot() +
  #geom_point() +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Number of Tools") +
  ylab("Proportion of Contigs in Common") +
  scale_fill_manual(name="",
                     values = alpha(pal(6), 0.3)) +
  scale_color_manual(name="",
                     values = alpha(pal(6), 0.7))

ggplot(prop_high_viruses_long[prop_high_viruses_long$vs==1,], aes(x=as.factor(numtools), y=proportion, 
                                  color=as.factor(numtools), fill=as.factor(numtools))) +
  geom_boxplot() +
  #geom_point() +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Number of Tools") +
  ylab("Proportion of Contigs in Common") +
  scale_fill_manual(name="",
                     values = alpha(pal(6), 0.3)) +
  scale_color_manual(name="",
                     values = alpha(pal(6), 0.7))

ggplot(prop_high_viruses_long[prop_high_viruses_long$vb==1,], aes(x=as.factor(numtools), y=proportion, 
                                  color=as.factor(numtools), fill=as.factor(numtools))) +
  geom_boxplot() +
  #geom_point() +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Number of Tools") +
  ylab("Proportion of Contigs in Common") +
  scale_fill_manual(name="",
                     values = alpha(pal(6), 0.3)) +
  scale_color_manual(name="",
                     values = alpha(pal(6), 0.7))

ggplot(prop_high_viruses_long[prop_high_viruses_long$dvf==1,], aes(x=as.factor(numtools), y=proportion, 
                                  color=as.factor(numtools), fill=as.factor(numtools))) +
  geom_boxplot() +
  #geom_point() +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Number of Tools") +
  ylab("Proportion of Contigs in Common") +
  scale_fill_manual(name="",
                     values = alpha(pal(6), 0.3)) +
  scale_color_manual(name="",
                     values = alpha(pal(6), 0.7))
```
rather than tool against all others, ones against ones, twos against twos...

```{r}
prop_high_viruses_long$numrules2 <- str_count(prop_high_viruses_long$tool, "1")

# accounting for rules that have multiple tools in them
prop_high_viruses_long$numtools2 <- prop_high_viruses_long$numrules2

prop_high_viruses_long$numtoolscommon <- 0

for (i in 1:nrow(prop_high_viruses_long)) {
  toolcombo <- as.character(prop_high_viruses_long$tool[i])
  if (substr(toolcombo, 1, 1)=="1") {
    prop_high_viruses_long$numtools2[i] <- prop_high_viruses_long$numtools2[i] + 3
  }

  if (substr(toolcombo, 5, 5)=="1") {
    prop_high_viruses_long$numtools2[i] <- prop_high_viruses_long$numtools2[i] + 4
  }

  if (prop_high_viruses_long$numtools2[i]==prop_high_viruses_long$numrules2[i]) {
    prop_high_viruses_long$numtoolscommon[i] <- prop_high_viruses_long$numtools2[i] 
  }

}

# can't actually have more than 6 tools
prop_high_viruses_long$numtools2[prop_high_viruses_long$numtools2>6] <- 6

prop_high_viruses_long$numrules2 <- as.factor(prop_high_viruses_long$numrules2)
prop_high_viruses_long$numtools2 <- as.factor(prop_high_viruses_long$numtools2)
prop_high_viruses_long$numtoolscommon <- as.factor(prop_high_viruses_long$numtoolscommon)
```

```{r}
ggplot(prop_high_viruses_long[prop_high_viruses_long$vs2==1,], aes(x=as.factor(numtoolscommon), y=proportion, 
                                  color=as.factor(numtoolscommon), fill=as.factor(numtoolscommon))) +
  geom_boxplot() +
  #geom_point() +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Number of Tools") +
  ylab("Proportion of Contigs in Common") +
  scale_fill_manual(name="",
                     values = alpha(pal(6), 0.3)) +
  scale_color_manual(name="",
                     values = alpha(pal(6), 0.7))

ggplot(prop_high_viruses_long[prop_high_viruses_long$vs==1,], aes(x=as.factor(numtoolscommon), y=proportion, 
                                  color=as.factor(numtoolscommon), fill=as.factor(numtoolscommon))) +
  geom_boxplot() +
  #geom_point() +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Number of Tools") +
  ylab("Proportion of Contigs in Common") +
  scale_fill_manual(name="",
                     values = alpha(pal(6), 0.3)) +
  scale_color_manual(name="",
                     values = alpha(pal(6), 0.7))

ggplot(prop_high_viruses_long[prop_high_viruses_long$vb==1,], aes(x=as.factor(numtoolscommon), y=proportion, 
                                  color=as.factor(numtoolscommon), fill=as.factor(numtoolscommon))) +
  geom_boxplot() +
  #geom_point() +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Number of Tools") +
  ylab("Proportion of Contigs in Common") +
  scale_fill_manual(name="",
                     values = alpha(pal(6), 0.3)) +
  scale_color_manual(name="",
                     values = alpha(pal(6), 0.7))

ggplot(prop_high_viruses_long[prop_high_viruses_long$dvf==1,], aes(x=as.factor(numtoolscommon), y=proportion, 
                                  color=as.factor(numtoolscommon), fill=as.factor(numtoolscommon))) +
  geom_boxplot() +
  #geom_point() +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Number of Tools") +
  ylab("Proportion of Contigs in Common") +
  scale_fill_manual(name="",
                     values = alpha(pal(6), 0.3)) +
  scale_color_manual(name="",
                     values = alpha(pal(6), 0.7))
```

# distribution of proportion in common for each pair

@@

```{r}
hist(prop_high_viruses_long$proportion)
```
